Title: Getting NJ assessment data into R: part 4 in a series.
Date: 2015-04-16
Status: draft
Category: education
Tags: NJ, assessment, NJASK, HSPA, data_management, tutorial
Slug: reading-nj-assess-data-4
Author: Andrew Martin

```{r, echo=FALSE}
#SET THIS TO TRUE WHEN READY TO PUBLISH
ready_to_ship = TRUE

library(knitr)
hook_plot <- knit_hooks$get('plot')

knit_hooks$set(plot=function(x, options) {
    if (!is.null(options$pelican.publish) && options$pelican.publish) {
        x <- paste0("{filename}", x)
    }
    hook_plot(x, options)
})
opts_chunk$set(dev='Cairo_svg')
opts_chunk$set(pelican.publish=ready_to_ship)

```

In my [last post]('/reading-nj-assess-data-3.html'), I showed how to extend the functions I developed for NJASK data to other state assessments.  In this post, I'll tie eveything together, and write some general functions that bring a wide variety of NJ state assessment data into R.

Roughly speaking, we're tying to write a function that will return data given a `year` and a `grade`.  Here are the big things that need to happen:

1. Check that call we made is a valid grade/year combination (raising an informative error if not)

2. Map the grade / year call to the correct `get_blank_data` function (NJASK? HSPA? GEPA?)

3. Fetch, clean, and return the data frame.

Let's tackle each of these in turn:

# valid calls

This function will test if a year/grade call is valid.

```{r valid_call}

valid_call <- function(year, grade) {
  #common core transition started in 2015
  if(year > 2014) {
    valid_call <- FALSE
  #NJASK fully implemented in 2008
  } else if(year >= 2006) {
    valid_call <- grade %in% c(3:8, 11)
  } else if (year >= 2004) {
    valid_call <- grade %in% c(3, 4, 8, 11)
  } else if (year < 2004) {
    valid_call <- FALSE
  }
  
  return(valid_call)

}

```

# map for retrieval

This function does normal retrieval (NJASK for 3-8; HSPA for 11).

```{r}

standard_assess <- function(year, grade) {
  if(grade %in% c(3:8)) {
    assess_data <- get_raw_njask(year, grade) %>%
      process_njask()
  } else if (grade == 11) {
    #not yet implemented
    assess_data <- get_raw_hspa(year) %>%
      process_hspa
  }
  
  return(assess_data)
} 

```

Here is a mapping function that calls the correct retrieval method, given grade and year.

```{r main_wrapper}

fetch_nj_assess <- function(year, grade) {

  is_valid <- valid_call(year, grade)
  
  if (year >= 2008) {
    assess_data <- standard_assess(year, grade)
  #2006 and 2007: NJASK 3rd-7th, GEPA 8th, HSPA 11th
  } else if (year %in% c(2006, 2007)) {
    if (grade %in% c(3:7)) {
      assess_data <- standard_assess(year, grade)  
    } else if (grade == 8) {
      #not yet implemented
      assess_data <- get_raw_gepa(year) %>%
        process_gepa()
    } else if (grade == 11) {
      #not yet implemented
      assess_data <- get_raw_hspa(year) %>%
        process_hspa
    }
  #2004 and 2005:  NJASK 3rd & 4th, GEPA 8th, HSPA 11th
  } else if (year %in% c(2004, 2005)) {
    if (grade %in% c(3:4)) {
      assess_data <- standard_assess(year, grade)  
    } else if (grade == 8) {
      #not yet implemented
      assess_data <- get_raw_gepa(year) %>%
        process_gepa()
    } else if (grade == 11) {
      #not yet implemented
      assess_data <- get_raw_hspa(year) %>%
        process_hspa 
    }
  # if we ever reached this block, there's a problem with our `valid_call()` function
  } else {
    #raise an error
  }
 
  return(assess_data)
}

fetch_nj_assess(2014, 6) %>% head()

```

# all together

Finally, as a convenience, let's write a function that brings down all of the NJASK data for all years and grades.

```{r all_njask}



```

