Title: data cleanup for longform post
Date: 2015-04-19
Category: data_prep
Status: draft
Slug: longform-data-cleanup
Author: Andrew Martin

Read in the file.

```{r raw_data}

lf <- read.csv('datasets/longform_sources.csv', stringsAsFactors=FALSE)

#html cleanup
lf$publication <- gsub('&#x27;', "'", lf$publication)

head(lf)

```

There are some magazine names in there that are obvious duplicates.  There are a whole host of de-duping/text processing methods that we could use here, but in this case I'm just going to fix the obvious ones and move on.

The most common dupes have the format 'X' and '_The_ X'.  We can look for those patterns in one go:

```{r the_x}

pubs <- unique(lf$publication)

dupes <- c()

for (i in pubs) {
  if (paste('The', i) %in% pubs) {
    dupes <- append(dupes, i)
  }
}

#print(dupes)
```

We can just loop over that list of dupe titles and make sure they all have the same name.

```{r fix_dupes}

for (j in dupes) {
  lf$publication <- gsub(paste('The', j), j, lf$publication)
}

```

Turn `posted_at` into a proper date:

```{r fix_date}
library(lubridate)
library(magrittr)

lf$posted_at <- lubridate::parse_date_time(
  lf$posted_at, orders=c('d-m-y','d-M-y')
)

```

Filter out bad records:

```{r}

lf <- lf %>%
  dplyr::filter(
    lubridate::year(lf$posted_at) >= 2010  
  )

```

Percentage share by year:

```{r by_year}
library(dplyr)

lf$year_posted <- lubridate::year(lf$posted_at)

lf_yr <- lf %>%
  group_by(year_posted) %>%
  summarize(
    n = n()
  )

names(lf_yr)[2] <- 'yr_total'

lf_pub_yr <- lf %>%
  group_by(publication, year_posted) %>%
  summarize(
    n_pub = n()  
  )

lf_pub_disp <- lf_pub_yr
lf_pub_disp$publication <- ifelse(
  (lf_pub_disp$year %in% c(2011, 2012, 2013, 2014) & lf_pub_disp$n < 5) | (lf_pub_disp$year %in% c(2010, 2015) & lf_pub_disp$n < 3), 'Other', lf_pub_disp$publication)

lf_pub_disp <- lf_pub_disp %>%
  group_by(publication, year_posted) %>%
  summarize(
    n_pub = sum(n_pub)
  ) %>%
  group_by(year_posted) %>%
  mutate(
    rank_pub = rank(desc(n_pub), ties.method='first')  
  )


```

Join back on year_posted

```{r join_back}

#all
lf_pub_yr <- dplyr::left_join(
  lf_pub_yr, lf_yr  
)

#display formatted
lf_pub_disp <- dplyr::left_join(
  lf_pub_disp, lf_yr  
)


lf_pub_yr$pct_year_total <- (lf_pub_yr$n_pub / lf_pub_yr$yr_total) * 100
lf_pub_disp$pct_year_total <- (lf_pub_disp$n_pub / lf_pub_disp$yr_total) * 100

head(lf_pub_yr)
head(lf_pub_disp)

```

